The CrypTool Script: \\ Mathematics and Cryptography \\ \vspace{30pt} {\large Background reading\\*[-8pt] for the free elearnig program CrypTool (version 1.3.10)} \vspace{60pt}

### 2005-07-04 script/en/chapters/digitalsignatures.tex
Update - Bernhard Esslinger June 2005

\newpage
% --------------------------------------------------------------------------
\section{Hash Functions and Digital Signatures}
\index{Signature!digital}
(Schneider J. / Esslinger B. / Filipovics B. / Koy H., June 2002,
Update: Feb. 2003)

@@

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} v1.3 you can also generate and check 
digital signatures: \\
using the submenus of the main menu {\bf Digital Signatures} or \\
using menu {\bf Indiv. Procedures \textbackslash{} RSA Demonstration 
\textbackslash{} Signature Demonstration}.
}.

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use 
a cryptographic hash function\footnote{%
Hash functions\index{Hash function} are implemented within 
CrypTool\index{CrypTool} v1.3.04 at several places.\\
Using menus {\bf Individual Procedures \textbackslash{} Hash} and
            {\bf Individual Procedures \textbackslash{} Attack on the Hash
	         Value of the Digital Signature}
you have the possibilities
% hier die items nicht einrücken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value and \\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
}.

% --------------------------------------------------------------------------
\subsection{Hash functions}
\index{Hash function}
A cryptographic \index{Hash function} {\em hash function}\footnote{%
Secure hash algorithms compute a condensed representation of 
electronic data (message).
When a message is input to a hash algorithm, the result is an output called
a message digest. The message digests typically range in length 
from 128 to 512 bits, depending on the algorithm. 
Secure hash algorithms are typically used with
other cryptographic algorithms, such as digital signature algorithms and
keyed-hash message authentication codes, or in the generation of random
numbers (bits).
}
maps a message of any
length to a string of characters with a constant size,
the \index{Hash value} hash value. It should be practically impossible, for a
given number, to find a message that has precisely this number as hash value.
Furthermore, it should be practically impossible to find two messages with the
same hash value\index{Collision}. 
In both cases the final signature procedure would have weaknesses.

So far, no formal proof has been found that perfectly secure cryptographic
hash functions exist. However, there are several good candidates that have
not yet shown any weak points in their structure in practice 
(e.g. \index{SHA-1} SHA--1\footnote{%
SHA-1 \index{SHA-1} is a 160 bit hash function specified in FIPS 180-1, 
ANSI X9.30 Part 2 and \cite{FIPS186}.\\
SHA means Secure Hash Algorithm, and is widely used, 
e.g. with DSA, RSA or ECDSA.\\
The current standard \cite{FIPS180-2} defines four secure hash algorithms
-- SHA-1, SHA-256, SHA-384, and SHA-512.
For these hash algorithms there are also validation tests defined in the
test suite FIPS 140-2.

The output length of the SHA algorithms was enhanced because of the
possibility of birthday attacks\index{Attack!birthday}\index{Collision}:
% \index{Birthday attack}
these make n-bit AES and a 2n-bit hash roughly equivalent: \\
- 128-bit AES -- SHA-256 \\
- 192-bit AES -- SHA-384 \\
- 256-bit AES -- SHA-512.

With CrypTool\index{CrypTool} v1.3.04 you can comprehend the birthday attack
\index{Attack!birthday} on digital signatures: \\
using menu {\bf Indiv. Procedures \textbackslash{} 
Attack on the hashvalue of the digital signature}.

}  or \index{RIPEMD-160} RIPEMD-160\footnote{%
RIPEMD-160, RIPEMD-128 and the optional extension RIPEMD-256 have object
identifiers defined by the ISO-identified organization TeleTrusT, both 
as hash algorithm and in combination with RSA.
RIPEMD-160 is also part of the ISO/IEC international standard 
ISO/IEC 10118-3:1998 on dedicated hash functions, together with 
RIPEMD-128 and SHA-1.

Further details: \\
http://www.esat.kuleuven.ac.be/~bosselae/ripemd160.html \\
http://www.ietf.org/rfc/rfc2857.txt (``The Use of HMAC-RIPEMD-160-96 within ESP and AH'').
}).

The hash function procedure is as follows:\\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

@@

% --------------------------------------------------------------------------
\subsection{RSA signatures}
\index{Signature!digital}
\index{RSA!signature}

@@

As explained above, signatures are not performed on the message itself, but
on a cryptographic hash value of the message. To prevent certain attacks
on the signature procedure (alone or in combination with encryption) it is
necessary to format the hash value before doing the exponentiation, as
described in the PKCS\#1 (Public Key Cryptography Standard \#1
\cite{PKCS1}). The fact that this standard had to be revised recently, after
being in use for several years, can serve as an example of how difficult it is
to get the details of cryptography right.

% --------------------------------------------------------------------------
\subsection{DSA signatures}
\index{Signature!digital}
\index{DSA!signature}

@@

% --------------------------------------------------------------------------
\subsection{Public key certification}
\index{Certification!public key}
\index{PKI}
The aim of public key certification is to guarantee the connection between a
public key and a user and to make it traceable for external parties. In cases in
which it is impossible to ensure that a public key really belongs to a
particular person, many protocols are no longer secure, even if the individual
cryptographic modules cannot be broken.

% --------------------------------------------------------------------------
\subsubsection{Impersonation attacks}\label{Impersonalisierungsattacke}
\index{Impersonation attack}
Assume Charlie has two pairs of keys (PK1, SK1) and (PK2, SK2), where SK denotes
the secret key and PK the public key. Further assume that he manages to palm off
PK1 on Alice as Bob's public key and PK2 on Bob as Alice's public key (by
falsifying a public key directory).

@@

% --------------------------------------------------------------------------
\subsubsection{X.509 certificate}
\index{X.509} 
Each participant who wants to have an X.509 certificate (\cite{X.509})
verifying that his public key belongs to a real person consults what is
known as a \index{Certification authority (CA)} certification authority
(CA)\footnote{%
Often called trust center, if the certificates are not only offered
to a closed user group.
}. 
He proves his identity to this CA (for example by showing his ID). 
The CA then issues him an electronic document (certificate) which 
essentially contains the name of the certificate-holder and the name
of the CA, the certificate-holder's public key and the validity period
of the certificate. The CA then signs the certificate using its secret key.

>>>>

\newpage
% --------------------------------------------------------------------------
\section{Hash Functions and Digital Signatures}
\hypertarget{Chapter_Hashes-and-Digital-Signatures}{}
\index{Signature!digital}
(Schneider J. / Esslinger B. / Koy H., June 2002,
Update: Feb. 2003, June 2005)

@@

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} you can also generate and check 
digital signatures: \\
using the submenus of the main menu {\bf Digital Signatures / PKI} or \\
using menu {\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem 
\textbackslash{} Signature Demonstration (Signature Generation)}.
}.

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. Because a signature is about
as long as the straight datastream to be signed, the procedure we just described
has in practice, however, a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we apply
a cryptographic hash function\footnote{%
Hash functions\index{Hash function} are implemented within 
CrypTool\index{CrypTool} at several places.\\
Using menus {\bf Individual Procedures \textbackslash{} Hash} and
            {\bf Analysis \textbackslash{} Hash}
you have the possibilities
% hier die items nicht einrücken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value,\\
\item to calculate a key from a password according to the PKCS\#5 
      standard\index{PKCS\#5}, \\
\item to calculate HMACs from a text and a secret key, and\\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
} to the document -- before signing. The output of the hash function will 
then be signed.

% --------------------------------------------------------------------------
\vskip + 15pt
\subsection{Hash functions}
\hypertarget{Hash-functions-ht}{}
\index{Hash function}
A {\em hash function}\footnote{%
Hash algorithms compute a condensed representation of electronic data (message).
When a message is input to a hash algorithm, the result is an output called
a message digest. The message digests typically range in length 
from 128 to 512 bits, depending on the algorithm. 
Secure hash algorithms are typically used with
other cryptographic algorithms, such as digital signature algorithms and
keyed-hash message authentication codes, or in the generation of random
numbers (bits).
}
maps a message of any length to a string of characters with a constant size,
the \index{Hash value} hash value.

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{Requirements for hash functions}

Cryptographically secure hash functions fulfill the following requirements:
\begin{itemize}
 \item Resistance against 1st Pre-Image attacks:
      \index{Pre-Image-Attack!1st}  \\
      It should be practically impossible, for a given number, to find a message
      that has precisely this number as hash value. \\
      Given (fix): hash value H', \\
      Searched: message m, so that: H(m) = H'.
 \item Resistance against 2nd Pre-Image attacks:
      \index{Pre-Image-Attack!2nd}  \\
      It should be practically impossible, for a given message, to find another
      message, which has precisely the same hash value. \\
      Given (fix): message m1 [and so the hash value H1 = H(m1)], \\
      Searched: message m2, so that: H(m2) = H1.
 \item Collision resistance:
      \index{Collision resistance}  \\
      It should be practically impossible to find any two messages with the
      same hash value (it doesn't matter what hash value). \\
      Searched: 2 messages m1 and m2, so that: H(m1) = H(m2).
\end{itemize}

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{Current attacks against hash functions like SHA-1}
\label{collision-attacks-against-sha-1}

So far, no formal proof has been found that perfectly secure cryptographic
hash functions exist.

During the past several years no new attacks against hash algorithms
came up, and so the candidates that had not yet shown any weaknesses
in their structure in practice 
(e.g. \index{SHA-1} SHA--1\footnote{%
  SHA-1 \index{SHA-1} is a 160 bit hash function specified in FIPS 180-1
  (by NIST), ANSI X9.30 Part 2 and \cite{FIPS186}.\\
  SHA means Secure Hash Algorithm, and is widely used, 
  e.g. with DSA, RSA or ECDSA.\\
  The current standard \cite{FIPS180-2} defines four secure hash algorithms
  -- SHA-1, SHA-256, SHA-384, and SHA-512.
  For these hash algorithms there are also validation tests defined in the
  test suite FIPS 140-2.

  The output length of the SHA algorithms was enhanced because of the
  possibility of birthday attacks\index{Attack!birthday}\index{Collision}:
  these make n-bit AES and a 2n-bit hash roughly equivalent: \\
  - 128-bit AES -- SHA-256 \\
  - 192-bit AES -- SHA-384 \\
  - 256-bit AES -- SHA-512.

  With CrypTool\index{CrypTool} you can comprehend the birthday attack
  \index{Attack!birthday} on digital signatures: \\
  using the menu {\bf Analysis \textbackslash{} Hash \textbackslash{} 
  Attack on the Hash Value of the Digital Signature}.
  }  
or \index{RIPEMD-160} RIPEMD-160\footnote{%
  RIPEMD-160, RIPEMD-128 and the optional extension RIPEMD-256 have object
  identifiers defined by the ISO-identified organization TeleTrusT, both 
  as hash algorithm and in combination with RSA.
  RIPEMD-160 is also part of the ISO/IEC international standard 
  ISO/IEC 10118-3:1998 on dedicated hash functions, together with 
  RIPEMD-128 and SHA-1.

  Further details: \\
  http://www.esat.kuleuven.ac.be/~bosselae/ripemd160.html \\
  http://www.ietf.org/rfc/rfc2857.txt (``The Use of HMAC-RIPEMD-160-96
  within ESP and AH'').
  }%
) were trusted.

At Crypto 2004 (August 2004)\footnote{%
    \href{http://www.iacr.org/conferences/crypto2004/}
 {\texttt{http://www.iacr.org/conferences/crypto2004/}} }
this safety-feeling was disputed: 
Chinese researchers published collision attacks against MD4, SHA-0 and
parts of SHA-1. This globally caused new motivation to engage in
new hash attack methods.

Right now the details of the results of the Chinese cryptographers are
only announced, but not revealed completely: They stated that collision
attacks on SHA-1 can be found with a work load of $2^{69}$.
This would mean that SHA-1 has cryptographic weaknesses, 
because the design of SHA-1 should ensure a work load for collision search
of about $2^{80}$.
The value of $2^{69}$ is currently a prognosis based on theoretic forecasts. 
The security of already created digital signatures is not affected by the
described attacks.

% be_2005_UPDATEN_if-hash-attacks-make-progress
According to our current knowledge there is no need to run scared.
But in the future digital signatures should use longer hash values and/or
other hash algorithms.

Already before Crypto 2004 NIST announced, to discontinue SHA-1 
in the next few years. So it is recommended not to use SHA-1 for
new products generating digital signatures.

Further information about this topic can be found in the article
``Hash cracked -- The consequences of the successful attacks on SHA-1''
by Reinhard Wobst and J\"urgen Schmidt\footnote{% 
      \href{http://www.heise.de/security/artikel/56634}
   {\texttt{http://www.heise.de/security/artikel/56634}}. \\
  Further references are e.g.: \\
      \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}
   {\texttt{http://www.bsi.bund.de/esig/basics/techbas/krypto/index.htm}} \\
      \href{http://csrc.nist.gov/CryptoToolkit/tkhash.html}
   {\texttt{http://csrc.nist.gov/CryptoToolkit/tkhash.html}}.
} 
  by Heise Security.

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{Signing with hash functions}
The hash function procedure is as follows:\\ Rather than signing the actual
document, the sender now first calculates the hash value of the message and
signs this. The recipient also calculates the hash value of the message (the
algorithm used must be known), then verifies whether the signature sent with the
message is a correct signature of the hash value. If this is the case, the
signature is verified to be correct. This means that the message is authentic,
because we have assumed that knowledge of the public key does not enable you to
derive the secret key. However, you would need this secret key to sign messages
in another name.

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsection{RSA signatures}
\index{Signature!digital}
\index{RSA!signature}

@@

As explained above, signatures are not performed on the message itself, but
on a cryptographic hash value of the message. To prevent certain attacks
on the signature procedure (alone or in combination with encryption) it is
necessary to format the hash value before doing the exponentiation, as
described in the PKCS\#1 (Public Key Cryptography Standard \#1
\cite{PKCS1})\index{PKCS\#1}. 
The fact that this standard had to be revised recently, after
being in use for several years, can serve as an example of how difficult it is
to get the details of cryptography right.

% --------------------------------------------------------------------------
\vskip + 15pt
\subsection{DSA signatures}
\index{Signature!digital}
\index{DSA!signature}

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsection{Public key certification}
\index{Certification!public key}
\index{PKI}
The aim of public key certification is to guarantee the connection between a
public key and a user and to make it traceable for external parties. In cases in
which it is impossible to ensure that a public key really belongs to a
particular person, many protocols are no longer secure, even if the individual
cryptographic modules cannot be broken.

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{Impersonation attacks}\label{Impersonalisierungsattacke}
\index{Impersonation attack}
Assume Charlie has two pairs of keys (PK1, SK1) and (PK2, SK2), where SK denotes
the secret key and PK the public key. Further assume that he manages to palm off
PK1 on Alice as Bob's public key and PK2 on Bob as Alice's public key (by
falsifying a public key directory).

@@

% --------------------------------------------------------------------------
\vskip + 15pt
\subsubsection{X.509 certificate}
\index{X.509} 
Each participant who wants to have an X.509 certificate (\cite{X.509})
verifying that his public key belongs to a real person consults what is
known as a \index{Certification authority (CA)} certification authority
(CA)\footnote{%
Often called trust center, if the certificates are not only offered
to a closed user group.
}. 
He proves his identity to this CA (for example by showing his ID). 
The CA then issues him an electronic document (certificate) which 
essentially contains the name of the certificate-holder and the name
of the CA, the certificate-holder's public key and the validity period
of the certificate. The CA then signs the certificate using its secret key.

### 2005-07-04 script/en/chapters/ellipticcurves.tex
be updates
*** empty log message ***
no message
Update - Bernhard Esslinger June 2005

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / B\"uger M. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001, June 2002, Mar. 2003)

\subsection{Elliptic curve cryptography -- a high-performance substitute for RSA?}\index{Performance}\label{ECAlternative}

@@

% -> Figure 1
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.75]{figures/RSAKeylength}
\caption{Prognosis of the key lengths to be regarded safe for RSA and
  Elliptic Curves\vspace{1ex}} 
\label{RSAKeylength}
\end{center}
\end{figure}

@@

% -> Figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.75]{figures/ThousandBitMultiplications}
\caption{Comparison of signing and verification time for RSA and Elliptic Curves} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}

@@

Mathematicians have been researching elliptic curves for over 100 years.
Over the course of time, many lengthy and mathematically complex results
have been found and published which are connected to elliptic curves. 
A mathematician would say that elliptic curves (or the mathematics behind
them) are widely understood. This research was originally purely 
mathematical. That is to say, elliptic curves were investigated, for 
example, in the mathematical areas of number theory and algebraic 
geometry, which are generally highly abstract. Even in the recent
past, elliptic curves played an important role in pure mathematics.
In 1993 and 1994, Andrew Wiles\index{Wiles Andrew} published 
mathematical works that triggered enthusiasm far beyond the specialist
audience. In these works, he proved a conjecture put forward in the 1960's.
To put it short, this conjecture was concerned with the connection 
between elliptic curves and what are called module forms. What is
particularly interesting for most people is that the works of Wiles
also proved the famous second theorem of Fermat. Mathematicians had
spent centuries (Fermat lived from 1601 to 1665) trying to find a
strict proof of this theorem. Understandably, therefore, Wiles' proof
got a good response. Fermat formulated his theorem as follows 
(written in the border of a book):

@@

Anyway that's enough about the role of elliptic curves in pure mathematics. 
In 1985 Neal Koblitz\index{Koblitz Neal} and Victor Miller\index{Miller Victor}
independently suggested using elliptic curves in cryptography. Elliptic 
curves have thus also found a concrete practical application. 
Another interesting area of application for elliptic curves is for
factorising whole numbers \index{Factorisation} (the RSA cryptographic system
is based on the \index{Complexity} difficulty/complexity of finding prime
factors of an extremely large number;  compare section \ref{SecurityRSA}.).
In this area, procedures based on elliptic curves have been investigated and
used since 1987 (compare section \ref{ECC-Factorisation}). \\
There are also prime number tests\index{Prime number!test} based on elliptic
curves.
% (s. Anmerkung U. Kuehn)

@@

There are factorisation%
\footnote{Especially John M. Pollard \index{Pollard John M.} was involved
in the development of many different factorisation algorithms; also at
factorisation with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contibutions in mathematics'':
\href{http://www.eff.org/Privacy/Crypto_misc/DESCracker/HTML/19990118_rsa_awards.html}
{\texttt{http://www.eff.org/Privacy/Crypto\_misc/DESCracker/HTML/19990118\_rsa\_awards.html}}.
}
algorithms based on elliptic curves%
\footnote{In 1987 H.W. Lenstra published
a factorization algorithm, based on elliptic curves (see \cite{Lenstra1987}).
The biggest compound number currently 
factorised\index{Factorisation!factoring records} with elliptic curves is 
the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi 
(See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

@@

CrypTool\index{CrypTool} also offers elliptic curves for the digital signature function.

@@

The procedure implements the cryptographic primitives for generating and
verifying signatures for the variations of Nyberg-Rueppel signatures and
\index{DSA} DSA signatures based on elliptic curves (in accordance with draft
no. 8 of the IEEE P1363 work group). This was done in collaboration with the
Secude GmbH --- using the above library and the Secude SDK.
\index{Secude}

@@

    \bibitem[Cassels1991]{Cassels1991} J. W. S. Cassels,
	\index{Cassels 1991} \\
        {\em Lectures on elliptic curves},\\
	Cambridge University Press, 1991, 143 pages.
    \bibitem[Koblitz1984]{Koblitz1984}N. Koblitz, 
	\index{Koblitz 1984} \\
        {\em Introduction to elliptic curves and modular forms},\\
	Graduate Texts in Mathemathics, Springer-Verlag, 1984.
    \bibitem[Koblitz1998]{Koblitz1998}N. Koblitz,  
	\index{Koblitz 1998}\\
	{\em Algebraic aspects of Cryptography. With an appendix on 
	Hyperelleptic curves by Alfred J. Menezes, Yi Hong Wu and Robert 
	J. Zuccherato}, \\
	Springer-Verlag, 1998, 206 pages.
    \bibitem[Menezes1993]{Menezes1993}A. J., Menezes, 
	\index{Menezes 1993} \\
        {\em Elliptic curve public key cryptosystems},\\
	Kluwer Academic Publishers, 1993.
    \bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra, 
        \index{Lenstra 1987} \\
        {\em Factoring integers with elliptic curves}, \\
	Annals of Mathematics 126, pp. 649-673, 1987.
    \bibitem[Lenstra1999]{Lenstra1999} Arjen K. Lenstra, Eric R. Verheul
        \index{Lenstra/Verheul 1999} \\
        {\em Selecting Cryptographic Key Sizes (1999)},\\
	Journal of Cryptology: the journal of the International 
	Association for Cryptologic Research \\
	\href{http://www.cryptosavvy.com/cryptosizes.pdf}
	{\texttt{http://www.cryptosavvy.com/cryptosizes.pdf}}
    \bibitem[Silverman1986]{Silverman1986} J. Silverman, 
        \index{Silverman 1986}\\
        {\em The Arithmetic of Elliptic Curves},\\
	Springer-Verlag, 1986.
    \bibitem[Silverman1992]{Silverman1992}J. Silverman,
	\index{Silverman 1992} \\
        {\em The arithmetic of elliptc curves},\\
	Graduate Texts in Mathemathics, Springer-Verlag, 1992.
    \bibitem[SivermanTate1992]{SivermanTate1992}J. Silverman, J. Tate,
	\index{Siverman/Tate 1992} \\
        {\em Rational points on elliptic curves},\\
	Springer-Verlag, 1992.

>>>>

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\label{Chapter_EllipticCurves}
\hypertarget{ellcurve}{}
(Filipovics B. / B\"uger M. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001, June 2002, Mar. 2003)

\subsection{Elliptic curve cryptography -- a high-performance substitute for RSA?}
\index{Performance}
\label{ECAlternative}

@@

% -> Figure 1
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.7]{figures/RSAKeyLength-2}
\caption{Prognosis of the key lengths to be regarded safe for RSA and
  Elliptic Curves\vspace{1ex}} 
\label{RSAKeylength}
\end{center}
\end{figure}

@@

% -> Figure 2
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.7]{figures/ECCRSA}
\caption{Comparison of signing and verification time for RSA and Elliptic Curves} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}

@@

Mathematicians have been researching elliptic curves for over 100 years.
Over the course of time, many lengthy and mathematically complex results
have been found and published which are connected to elliptic curves. 
A mathematician would say that elliptic curves (or the mathematics behind
them) are widely understood. This research was originally purely 
mathematical. That is to say, elliptic curves were investigated, for 
example, in the mathematical areas of number theory and algebraic 
geometry, which are generally highly abstract. Even in the recent
past, elliptic curves played an important role in pure mathematics.
In 1993 and 1994, Andrew Wiles\index{Wiles, Andrew} published 
mathematical works that triggered enthusiasm far beyond the specialist
audience. In these works, he proved a conjecture put forward in the 1960's.
To put it short, this conjecture was concerned with the connection 
between elliptic curves and what are called module forms. What is
particularly interesting for most people is that the works of Wiles
also proved the famous second theorem of Fermat. Mathematicians had
spent centuries (Fermat lived from 1601 to 1665) trying to find a
strict proof of this theorem. Understandably, therefore, Wiles' proof
got a good response. Fermat formulated his theorem as follows 
(written in the border of a book):

@@

Anyway that's enough about the role of elliptic curves in pure mathematics. 
In 1985 Neal Koblitz\index{Koblitz, Neal} and Victor Miller\index{Miller, Victor}
independently suggested using elliptic curves in cryptography. Elliptic 
curves have thus also found a concrete practical application. 
Another interesting area of application for elliptic curves is for
factorising whole numbers \index{Factorisation} (the RSA cryptographic system
is based on the \index{Complexity} difficulty/complexity of finding prime
factors of an extremely large number;  compare section \ref{SecurityRSA}.).
In this area, procedures based on elliptic curves have been investigated and
used since 1987 (compare section \ref{ECC-Factorisation}). \\
There are also prime number tests\index{Prime number!test} based on elliptic
curves.
% (s. Anmerkung U. Kuehn)

@@

There are factorisation%
\footnote{Especially John M. Pollard \index{Pollard, John M.} was involved
in the development of many different factorisation algorithms; also at
factorisation with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contributions in mathematics'':
\href{http://www.eff.org/Privacy/Crypto_misc/DESCracker/HTML/19990118_rsa_awards.html}
{\texttt{http://www.eff.org/Privacy/Crypto\_misc/DESCracker/HTML/19990118\_rsa\_awards.html}}.
}
algorithms based on elliptic curves%
\footnote{In 1987 H.W. Lenstra published
a factorization algorithm, based on elliptic curves (see \cite{Lenstra1987}).
The biggest compound number currently 
factorised\index{Factorisation!factoring records} with elliptic curves is 
the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi 
(See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

@@

CrypTool\index{CrypTool} also offers elliptic curves for the digital signature function%
\footnote{%
The dialog box, which appears in CrypTool\index{CrypTool} after clicking the menu
{\bf Digital Signatures/PKI \textbackslash{} Sign Message},
offers the EC methods ECSP-DSA and ECSP-NR.
}.

@@

The procedure implements the cryptographic primitives for generating and
verifying signatures for the variations of Nyberg-Rueppel signatures and
\index{DSA} DSA signatures based on elliptic curves (in accordance with draft
no. 8 of the IEEE P1363 work group). This was done in collaboration with the
Secude GmbH --- using the above library and the Secude SDK.
\index{SECUDE IT Security}

@@

    \bibitem[Cassels1991]{Cassels1991} J. W. S. Cassels,
	\index{Cassels 1991} \\
        {\em Lectures on elliptic curves},\\
	Cambridge University Press, 1991, 143 pages.
    \bibitem[Koblitz1984]{Koblitz1984}N. Koblitz, 
	\index{Koblitz 1984} \\
        {\em Introduction to elliptic curves and modular forms},\\
	Graduate Texts in Mathemathics, Springer-Verlag, 1984.
    \bibitem[Koblitz1998]{Koblitz1998}N. Koblitz,  
	\index{Koblitz 1998}\\
	{\em Algebraic aspects of Cryptography. With an appendix on 
	Hyperelleptic curves by Alfred J. Menezes, Yi Hong Wu and Robert 
	J. Zuccherato}, \\
	Springer-Verlag, 1998, 206 pages.
    \bibitem[Menezes1993]{Menezes1993}A. J. Menezes, 
	\index{Menezes 1993} \\
        {\em Elliptic curve public key cryptosystems},\\
	Kluwer Academic Publishers, 1993.
    \bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra, 
        \index{Lenstra 1987} \\
        {\em Factoring integers with elliptic curves}, \\
	Annals of Mathematics 126, pp. 649-673, 1987.
    \bibitem[Lenstra1999]{Lenstra1999} Arjen K. Lenstra, Eric R. Verheul
        \index{Lenstra/Verheul 1999} \\
        {\em Selecting Cryptographic Key Sizes (1999)},\\
	Journal of Cryptology: the journal of the International 
	Association for Cryptologic Research \\
	\href{http://www.cryptosavvy.com/cryptosizes.pdf}
	{\texttt{http://www.cryptosavvy.com/cryptosizes.pdf}}
    \bibitem[Silverman1986]{Silverman1986} J. Silverman, 
        \index{Silverman 1986}\\
        {\em The Arithmetic of Elliptic Curves},\\
	Springer-Verlag, 1986.
    \bibitem[Silverman1992]{Silverman1992}J. Silverman,
	\index{Silverman 1992} \\
        {\em The arithmetic of elliptc curves},\\
	Graduate Texts in Mathemathics, Springer-Verlag, 1992.
    \bibitem[SilvermanTate1992]{SilvermanTate1992}J. Silverman, J. Tate,
	\index{Siverman/Tate 1992} \\
        {\em Rational points on elliptic curves},\\
	Springer-Verlag, 1992.
